---
title: "Make a secure shipment"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(sship)
```

This documents describes how to make secure shipments with the R package _sship_. While _sship_ provides technical means of very strong protections throug encryption of digital content all still depends on proper and knowledgeable use. Therefore, please take a moment and study the next paragraph very carefully.

_sship_ ***makes sure that a secret will only be available to the recipient by applying the recipent's public key. When*** _sship_ ***is to be used for sending a secret to a recipent one must be absouletly sure that the public key used by*** _sship_ ***belongs to the intended recipient. The authenticity of the public key should be established by a combination of technical and social means. Techically, the public key should be managed (by its owner, the recipient) throug a key provider that enforces multi-factor authentication of its users. Socially, the two parties to share a secret should know of each other (personally, profesionally or similar) and establish trust both between themselves and the technology to be used ensure authenticity of the public key.***

Following is an example of how the authenticity of a public key may be sufficiently established. Person or entity X needs to send a secret data file to person or entity Y and therefor need Y's public key. A public key provider exists (_e.g._ GitHub) and within this provider X has made a dedicated group (or organization in GitHub-terms) that enforces multi-factor authentication of its members. X asks Y to apply for a membership of this group at the public key provider. X now expects a member request from Y and grants it as it arrives. Since multi-factor authentication is enforced it is unlikely that anyone else but Y may add, alter or delete Y's public key at the provider. X can thereafter collect Y's public key from the provider at any time and be sufficiently sure that it will be authentic. 
